Learning and comparing functional connectomes across subjects

### 2013-03-26 paper.tex
Make the paper suitable for compilation in postscript
A few typos

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Neuroimage-like layout
\documentclass[5p]{elsarticle}
% For kindle
%\documentclass[1p,12pt]{elsarticle}
%\usepackage{geometry}
%\geometry{a6paper,hmargin={.2cm,.2cm},vmargin={1cm,1cm}}
% End kindle
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{bm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}
\usepackage[breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
\usepackage[table]{xcolor}

@@

@@

%{{{------ Parcellation figure -----------------------------------------------
\begin{figure}
\includegraphics[width=.5\linewidth]{aal_atlas}%
\hfill%
\includegraphics[width=.5\linewidth]{ho_atlas}%

@@

\includegraphics[width=.5\linewidth]{smith_atlas}%
\hfill%
\includegraphics[width=.5\linewidth]{msdl_atlas}%

@@

%{{{------ Correlation matrices figure ---------------------------------------
\begin{figure}
\newcommand{\photoframe}[1]{\setlength\fboxsep{0pt}%
{\color{black!70}\fbox{#1}\color{black}}}%
\begin{minipage}{.666\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_cov}%
    \raisebox{.606\linewidth}{%
	\hspace*{-.462\linewidth}%
	\photoframe{%
	%\setlength\fboxsep{1pt}%
	\colorbox{white}{%
	\raisebox{.003\linewidth}{%
	    \includegraphics[width=.455\linewidth]{group_emp_cov_3d_labels}}%
	\rule{0pt}{.387\linewidth}%
	}}%
    }
\end{minipage}%
\hspace*{1pt}%
\begin{minipage}{.33\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_cov_no_confounds}%

    \includegraphics[width=\linewidth]{group_emp_cov_global_mean}%
\end{minipage}%

@@

%{{{------ Precision matrices figure -----------------------------------------
\begin{figure}
\newcommand{\photoframe}[1]{\setlength\fboxsep{0pt}%
{\color{black!70}\fbox{#1}\color{black}}}%
\begin{minipage}{.666\linewidth}%
    \includegraphics[width=\linewidth]{group_l21_prec}%
    \raisebox{.606\linewidth}{%
	\hspace*{-.462\linewidth}%
	\photoframe{%
	%\setlength\fboxsep{1pt}%
	\colorbox{white}{%
	\raisebox{.003\linewidth}{%
	    \includegraphics[width=.455\linewidth]{group_l21_prec_3d_labels}}%
	\rule{0pt}{.387\linewidth}%
	}}%
    }
\end{minipage}%
\hspace*{1pt}%
\begin{minipage}{.33\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_prec}%

    \includegraphics[width=\linewidth]{group_l1_prec}%
\end{minipage}%

@@

%{{{------ Regressor in precision matrix -------------------------------------
\begin{figure}
\includegraphics[height=.499\linewidth]{group_l21_prec_no_confounds}%
\hfill%
\includegraphics[height=.499\linewidth]{group_l21_prec_global_mean}%

@@

%{{{------ Inter subject fluctuations ----------------------------------------
\begin{figure}
\hspace*{.5ex}%
\includegraphics[width=.196\linewidth]{inter_subj/cov_sub00}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/cov_sub29}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/cov_sub01}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/cov_sub07}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/cov_sub53}%
\hspace*{.1ex}%
\raisebox{.005\linewidth}{%
    \includegraphics[width=.0186\linewidth]{cmap_00}%
    \llap{
	\raisebox{.003\linewidth}{%
	    \bfseries\tiny\color{white} -\!1\hspace*{-.2ex}}%
    }%
    \llap{
	\raisebox{.17\linewidth}{%
	    \bfseries\tiny\color{white} 1\hspace*{.1ex}}%
    }%
}%
\llap{\rlap{
    \setlength\fboxsep{1pt}%
    \raisebox{.16\linewidth}{%
	\colorbox{black!10}{%
	    \sffamily\small\textbf{a}. Correlation matrices}}%
	}%
    \hspace*{.993\linewidth}%
}

@@

\hspace*{.5ex}%
\includegraphics[width=.196\linewidth]{inter_subj/res_sub00}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/res_sub29}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/res_sub01}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/res_sub07}%
\hspace*{-.2ex}%
\includegraphics[width=.196\linewidth]{inter_subj/res_sub53}%
\hspace*{.1ex}%
\raisebox{.005\linewidth}{%
    \includegraphics[width=.0186\linewidth]{cmap_01}%
    \llap{
	\raisebox{.003\linewidth}{%
	    \bfseries\tiny -\!\hspace*{-.1ex}4\hspace*{-.1ex}}%
    }%
    \llap{
	\raisebox{.168\linewidth}{%
	    \bfseries\tiny 4\hspace*{.1ex}}%
    }%
}%
\llap{\rlap{
    \setlength\fboxsep{1pt}%
    \raisebox{.16\linewidth}{%
	\colorbox{black!10}{%
	    \sffamily\small\textbf{e}. Z score on residuals \scriptsize
		\raisebox{.2ex}{\cite{varoquaux2010b}}}}%
	}%
    \hspace*{.993\linewidth}%
}

@@

%{{{------ Integration measure -----------------------------------------------
\begin{figure}
\hspace*{2ex}%
\begin{minipage}{.19\linewidth}
    \center\sffamily
    {\small Integration:}

@@

    2.422
\end{minipage}%
\hspace*{2ex}%

@@

\paragraph{Graph-topological metrics}
%
Functional connectivity graphs have been found to display specific
topological properties\footnote{In the neuroscience world, these
descriptions are grouped under the terms of ``graph-theoretical
approaches'', however graph theory is an entire division of mathematics
and computer science that is concerned with much more than topology of
random graphs.} that are characteristic of small-word networks
\cite{stam2004,salvador2005,achard2006,bullmore2009}. These networks
display excellent transport properties: although they have a relatively
small number of connections, any two regions of the brain are well
connected. 
Another interesting consequence of their specific topology is
the resilience it gives the system to attacks such as resulting from
brain lesions \cite{achard2006}. This overall structure of
functional-connectivity graphs can be summarized by a few metrics, such
as the average path length between any two nodes, the local clustering
coefficients, or the node degree centrality \cite{rubinov2010}. Given that pathologies without a
localized focus, such
as schizophrenia, are thought to have a global impact on brain
connectivity \cite{liu2008,bassett2008}, the graph-topological metrics are
promising markers to perform inter-subject comparison. Such an
approach is appealing as it is not subject to multiple comparison issues.
However, it has been criticized as giving a fairly unspecific
characterization of the brain and being fragile to noise
\cite{ioannides2007}. Another caveat is that these properties are not
specific to brain function: correlation matrices display
small-word properties such as local clustering by construction. Indeed, if two
nodes are strongly correlated to a third, they are highly likely to be
correlated to each other \cite{zalesky2012}. This observation highlights
the need for well defined null-hypothesis \cite{zalesky2012,rubinov2011},
but also for controlled recovery of brain functional connectivity going
beyond empirical correlation matrices, as discuss in the previous section.

@@

Predictive modeling is concerned with learning (or fitting) a model that is
capable of \emph{predicting} information from unseen data
\cite{pereira2009}.  In the context of connectomes, predictive modeling
can extract connectivity-based biomarkers of disease diagnosis,
prognosis, or other phenotypic outcomes \cite{craddock2009, dosenbach2010}. The
accuracy of a predictive model provides a measure of the amount of information
present in the connectome about the phenotypic measure being evaluated
\cite{kjems2002, kriegeskorte2006}.  When combined with reproducibility, prediction accuracy
provides a metric for evaluating experimental trade-offs for data acquisition,
preprocessing, and analysis \cite{strother2002, laconte2003}.  Multivariate
predictive models are attractive in connectomics because they are sensitive to dependencies
between features and avoid the need to correct for multiple comparisons since
the significance of an entire pattern is evaluated using a single statistical
test. 
Additionally modern predictive modeling techniques draw from the
statistical learning literature, which specifically addresses
 high dimensional datasets with few observations.
Predictive modeling has been
successfully applied to identify connectome-based biomarkers of Alzheimer's
disease \cite{stonnington2010}, depression \cite{craddock2009, zeng2012}, schizophrenia
\cite{cecchi2009, shen2010}, autism \cite{anderson2011}, ADHD \cite{zhu2008}, aging
\cite{dosenbach2010}, as well as to classify mental operations
\cite{richiardi2011,shirer2012}. The growing interest for applying
predictive modeling to connectivity analysis was highlighted by the ADHD200
Global Competition, in which the object was to identify a connectivity-based
biomarker of ADHD \cite{adhd2002012}. Recent work has illustrated the utility
of predictive modeling for deriving connectivity models at the individual level
\cite{chu2011}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Neuroimage-like layout
\documentclass[5p]{elsarticle}
% For kindle
%\documentclass[1p,12pt]{elsarticle}
%\usepackage{geometry}
%\geometry{a6paper,hmargin={.2cm,.2cm},vmargin={1cm,1cm}}
% End kindle
\usepackage{graphicx}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{bm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}
\usepackage[breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
\usepackage[table]{xcolor}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% For the final version: to output PDF figures from latex
\newif\iffinal
\finaltrue
%\finalfalse
\iffinal\else
\usepackage[tightpage,active]{preview}
\fi

@@

%{{{------ Parcellation figure -----------------------------------------------
\begin{figure}
\iffinal
    % Final version
    \includegraphics[width=\linewidth]{pg_0001}%
\else
    % Commented on figure generation
    \begin{preview}
    \includegraphics[width=.5\linewidth]{aal_atlas}%
    \hfill%
    \includegraphics[width=.5\linewidth]{ho_atlas}%

@@

    \includegraphics[width=.5\linewidth]{smith_atlas}%
    \hfill%
    \includegraphics[width=.5\linewidth]{msdl_atlas}%
    \end{preview}
\fi

@@

%{{{------ Correlation matrices figure ---------------------------------------
\begin{figure}
\iffinal
    % Final version
    \includegraphics[width=\linewidth]{pg_0002}%
\else
    % Commented out figure generation
    \begin{preview}
    \newcommand{\photoframe}[1]{\setlength\fboxsep{0pt}%
    {\color{black!70}\fbox{#1}\color{black}}}%
    \begin{minipage}{.666\linewidth}%
	\includegraphics[width=\linewidth]{group_emp_cov}%
	\raisebox{.606\linewidth}{%
	    \hspace*{-.462\linewidth}%
	    \photoframe{%
	    %\setlength\fboxsep{1pt}%
	    \colorbox{white}{%
	    \raisebox{.003\linewidth}{%
	       \includegraphics[width=.455\linewidth]{group_emp_cov_3d_labels}}%
	    \rule{0pt}{.387\linewidth}%
	    }}%
	}
    \end{minipage}%
    \hspace*{1pt}%
    \begin{minipage}{.33\linewidth}%
	\includegraphics[width=\linewidth]{group_emp_cov_no_confounds}%

	\includegraphics[width=\linewidth]{group_emp_cov_global_mean}%
    \end{minipage}%
    \end{preview}
\fi

@@

%{{{------ Precision matrices figure -----------------------------------------
\begin{figure}
\iffinal
    % Final version
    \includegraphics[width=\linewidth]{pg_0003}%
\else
    % Commented out figure generation
    \begin{preview}
    \newcommand{\photoframe}[1]{\setlength\fboxsep{0pt}%
    {\color{black!70}\fbox{#1}\color{black}}}%
    \begin{minipage}{.666\linewidth}%
	\includegraphics[width=\linewidth]{group_l21_prec}%
	\raisebox{.606\linewidth}{%
	    \hspace*{-.462\linewidth}%
	    \photoframe{%
	    %\setlength\fboxsep{1pt}%
	    \colorbox{white}{%
	    \raisebox{.003\linewidth}{%
	      \includegraphics[width=.455\linewidth]{group_l21_prec_3d_labels}}%
	    \rule{0pt}{.387\linewidth}%
	    }}%
	}
    \end{minipage}%
    \hspace*{1pt}%
    \begin{minipage}{.33\linewidth}%
	\includegraphics[width=\linewidth]{group_emp_prec}%

	\includegraphics[width=\linewidth]{group_l1_prec}%
    \end{minipage}%
    \end{preview}
\fi

@@

%{{{------ Regressor in precision matrix -------------------------------------
\begin{figure}
\iffinal
    % Final version
    \includegraphics[width=\linewidth]{pg_0004}%
\else
    % Commented out figure generation
    \begin{preview}
    \includegraphics[height=.499\linewidth]{group_l21_prec_no_confounds}%
    \hfill%
    \includegraphics[height=.499\linewidth]{group_l21_prec_global_mean}%
    \end{preview}
\fi

@@

%{{{------ Inter subject fluctuations ----------------------------------------
\begin{figure}
\iffinal
    % Final version
    \includegraphics[width=\linewidth]{pg_0005}%
\else
    % Commented out figure generation
    \begin{preview}
    \hspace*{.5ex}%
    \includegraphics[width=.196\linewidth]{inter_subj/cov_sub00}%
    \hspace*{-.2ex}%
    \includegraphics[width=.196\linewidth]{inter_subj/cov_sub29}%
    \hspace*{-.2ex}%
    \includegraphics[width=.196\linewidth]{inter_subj/cov_sub01}%
    \hspace*{-.2ex}%
    \includegraphics[width=.196\linewidth]{inter_subj/cov_sub07}%
    \hspace*{-.2ex}%
    \includegraphics[width=.196\linewidth]{inter_subj/cov_sub53}%
    \hspace*{.1ex}%
    \raisebox{.005\linewidth}{%
	\includegraphics[width=.0186\linewidth]{cmap_00}%
	\llap{
	    \raisebox{.003\linewidth}{%
		\bfseries\tiny\color{white} -\!1\hspace*{-.2ex}}%
	}%
	\llap{
	    \raisebox{.17\linewidth}{%
		\bfseries\tiny\color{white} 1\hspace*{.1ex}}%
	}%
    }%
    \llap{\rlap{
	\setlength\fboxsep{1pt}%
	\raisebox{.16\linewidth}{%
	    \colorbox{black!10}{%
		\sffamily\small\textbf{a}. Correlation matrices}}%
	    }%
	\hspace*{.993\linewidth}%
    }

@@

    \hspace*{.5ex}%
    \includegraphics[width=.196\linewidth]{inter_subj/res_sub00}%
    \hspace*{-.2ex}%
    \includegraphics[width=.196\linewidth]{inter_subj/res_sub29}%
    \hspace*{-.2ex}%
    \includegraphics[width=.196\linewidth]{inter_subj/res_sub01}%
    \hspace*{-.2ex}%
    \includegraphics[width=.196\linewidth]{inter_subj/res_sub07}%
    \hspace*{-.2ex}%
    \includegraphics[width=.196\linewidth]{inter_subj/res_sub53}%
    \hspace*{.1ex}%
    \raisebox{.005\linewidth}{%
	\includegraphics[width=.0186\linewidth]{cmap_01}%
	\llap{
	    \raisebox{.003\linewidth}{%
		\bfseries\tiny -\!\hspace*{-.1ex}4\hspace*{-.1ex}}%
	}%
	\llap{
	    \raisebox{.168\linewidth}{%
		\bfseries\tiny 4\hspace*{.1ex}}%
	}%
    }%
    \llap{\rlap{
	\setlength\fboxsep{1pt}%
	\raisebox{.16\linewidth}{%
	    \colorbox{black!10}{%
		\sffamily\small\textbf{e}. Z score on residuals \scriptsize
		    \raisebox{.2ex}{\cite{varoquaux2010b}}}}%
	    }%
	\hspace*{.993\linewidth}%
    }
    \end{preview}
\fi

@@

%{{{------ Integration measure -----------------------------------------------
\begin{figure}
\iffinal
    % Final version
    \includegraphics[width=\linewidth]{pg_0006}
\else
    % Commented out figure generation
    \begin{preview}
    \hspace*{2ex}%
    \begin{minipage}{.19\linewidth}
	\center\sffamily
	{\small Integration:}

@@

	2.422
    \end{minipage}%
    \hspace*{2ex}%
    \end{preview}
\fi

@@

\paragraph{Graph-topological metrics}
%
Functional connectivity graphs have been found to display specific
topological properties\footnote{In the neuroscience world, these
descriptions are grouped under the terms of ``graph-theoretical
approaches'', however graph theory is an entire division of mathematics
and computer science that is concerned with much more than topology of
random graphs.} that are characteristic of small-world networks
\cite{stam2004,salvador2005,achard2006,bullmore2009}. These networks
display excellent transport properties: although they have a relatively
small number of connections, any two regions of the brain are well
connected. 
Another interesting consequence of their specific topology is
the resilience it gives the system to attacks such as resulting from
brain lesions \cite{achard2006}. This overall structure of
functional-connectivity graphs can be summarized by a few metrics, such
as the average path length between any two nodes, the local clustering
coefficients, or the node degree centrality \cite{rubinov2010}. Given that pathologies without a
localized focus, such
as schizophrenia, are thought to have a global impact on brain
connectivity \cite{liu2008,bassett2008}, the graph-topological metrics are
promising markers to perform inter-subject comparison. Such an
approach is appealing as it is not subject to multiple comparison issues.
However, it has been criticized as giving a fairly unspecific
characterization of the brain and being fragile to noise
\cite{ioannides2007}. Another caveat is that these properties are not
specific to brain function: correlation matrices display
small-world properties such as local clustering by construction. Indeed, if two
nodes are strongly correlated to a third, they are highly likely to be
correlated to each other \cite{zalesky2012}. This observation highlights
the need for well defined null-hypothesis \cite{zalesky2012,rubinov2011},
but also for controlled recovery of brain functional connectivity going
beyond empirical correlation matrices, as discussed in the previous section.

@@

Predictive modeling is concerned with learning (or fitting) a model that is
capable of \emph{predicting} information from unseen data
\cite{pereira2009}.  In the context of connectomes, predictive modeling
can extract connectivity-based biomarkers of disease diagnosis,
prognosis, or other phenotypic outcomes \cite{craddock2009, dosenbach2010}. The
accuracy of a predictive model provides a measure of the amount of information
present in the connectome about the phenotypic measure being evaluated
\cite{kjems2002, kriegeskorte2006}.  When combined with reproducibility, prediction accuracy
provides a metric for evaluating experimental trade-offs for data acquisition,
preprocessing, and analysis \cite{strother2002, laconte2003}.  Multivariate
predictive models are attractive in connectomics because they are sensitive to dependencies
between features and avoid the need to correct for multiple comparisons since
the significance of an entire pattern is evaluated using a single statistical
test. 
Additionally, modern predictive modeling techniques draw from the
statistical learning literature, which specifically addresses
 high dimensional datasets with few observations.
Predictive modeling has been
successfully applied to identify connectome-based biomarkers of Alzheimer's
disease \cite{stonnington2010}, depression \cite{craddock2009, zeng2012}, schizophrenia
\cite{cecchi2009, shen2010}, autism \cite{anderson2011}, ADHD \cite{zhu2008}, aging
\cite{dosenbach2010}, as well as to classify mental operations
\cite{richiardi2011,shirer2012}. The growing interest for applying
predictive modeling to connectivity analysis was highlighted by the ADHD200
Global Competition, in which the object was to identify a connectivity-based
biomarker of ADHD \cite{adhd2002012}. Recent work has illustrated the utility
of predictive modeling for deriving connectivity models at the individual level
\cite{chu2011}.
